# Results {#results}

## Choice Analysis

We estimated the daily activity pattern choice models using mlogit for R
[@R-base, @mlogit2020]. As described above, the alternatives for daily activity
pattern choice are a Mandatory pattern where the individual's day involves a
work or school tour, a Non-Mandatory pattern where only discretionary trips are
taken, and a Home pattern where the individual does not leave home during the
day. In the models estimated for this study, the Home pattern serves as the
reference alternative with a utility of zero. Retired and otherwise non-working
individuals choose only between Non-Mandatory and Home daily activity patterns.

The model estimates are presented in 
Table \@ref(tab:model-coef). 
The estimated coefficients are of the expected sign, though not all are significant. 
Some predictors that proved to be insignificant, such as automobile availability
for full-time workers, were excluded from the estimated models. 
The overall model fit --- as indicated by the McFadden $\rho^2$ with respect to
a market shares (constants only) model --- is not strikingly high. Were the purpose
of this research to identify the best fit model of activity pattern choice for 
each person segment we would undertake an exercise to include, exclude, and identify
potential transformations for different sets of variables. In this case, however,
the goal of these models is simply to provide a plausible comparison point for
the behavior of individuals using wheelchairs against the behavior of individuals
in other person type segments. 

```{r model-coef}
tar_load(pt_modelsummary)

f <- function(x) format(round(x, 3))
gm <- list(
  list("raw" = "nobs", "clean" = "N", "fmt" = f),
  list("raw" = "AIC", "clean" = "AIC", "fmt" = f),
  list("raw" = "logLik", "clean" = "Log likelihood", "fmt" = f),
  list("raw" = "rho2", "clean" = "$\\rho^2$", "fmt" = f))
modelsummary(
  pt_modelsummary, group = variable + term ~ model, 
  gof_map = gm,
  coef_map = c(
    "M" = "Mandatory",
    "NM" = "Non-mandatory"
  ),
  estimate = "{estimate} ({statistic}){stars}",
  statistic = NULL,
  escape = FALSE,
  stars = c('*' = .01, '**' = .01),
  notes = c("t-statistics in parentheses, * p < 0.5, ** p < 0.01",
            "Coefficients represent utility change relative to stay at home pattern."),
  title = "Daily Activity Pattern Model Estimates") %>%
  kable_styling(latex_options = c("scale_down"))
```

The model coefficients indicate expected behavior for most included variables. 
All else equal, the model intercepts imply that full- and part-time workers 
are more likely to choose a mandatory daily pattern than either a non-mandatory 
home pattern. Men of all person types are less likely to choose non-mandatory 
patterns, and college graduates are more likely to choose any kind of out-of-home
pattern. The effect of age group and income are less impactful for most trip
purposes, with the exception of non-mandatory patterns being more likely to be
chosen by individuals between the ages of 65 and 79. The strong coefficient 
seen on the choices of full-time workers over 79 years old is not significant, 
and is rather a relic of very few full-time workers of that age in the data set.

The use of a wheelchair is strongly significant for all person segments, and
indicates wheelchair users are substantially less likely to make mandatory tours
even if employed. For non-workers and retirees who use wheelchairs, their propensity
to make non-mandatory tours is also substantially diminished. As a note, we considered
making wheelchair use an independent person type segment; the results suggest that 
full-time workers who use wheelchairs are more similar to other full-time workers
than they are to non-workers who also use wheelchairs. This justifies including
wheelchair use as a variable within each person type segment. More notable than
the significance of this variable alone, however, is the fact that its magnitude
is so strong. Indeed, wheelchair use appears to be more influential on the choice
of daily pattern than any other variable included in these models.


## Activity-based Model

At the individual level, wheelchair use appears highly predictive. It still remains
to be seen, however, what the effect of including this variable in a forecasting
model will do to aggregate trip making. To examine this question, we placed the 
wheelchair use variable coefficients shown in Table \@ref(tab:model-coef) into 
the ActivitySim daily activity pattern model. We allowed the other coefficients
to retain their values as originally specified, assuming that wheelchair use 
is orthogonal to the other variables.


```{r dap-summary}
tar_load(dap_table)
kbl(dap_table, booktabs = TRUE, caption = "Daily Activity Pattern Change",
    col.names = c("Group", "Base", "Home", "Mandatory", "Non-Mandatory")) %>%
  add_header_above(c("", "", "Considering Wheelchair Use" = 3)) %>%
  collapse_rows(columns = 1, latex_hline = "major") %>% 
  kable_styling()
```

```{r dap-calcs}
tar_load(asim_dap)
n_changed <- nrow(asim_dap %>% filter(DAP_after != DAP_before))
n_total <- nrow(asim_dap)
change1 <- (932 + 308) / (932 + 1642 + 308)
change2 <- (1235 + 759) / (1235 + 759 + 4511)
```

Overall, `r n_changed` individuals of the `r n_total` 
in the region changed their daily activity patterns changed from the base
scenario after including wheelchair use as a utility variable.
Though this is only `r n_changed / n_total * 100` percent of individuals, it is worth
considering the distribution of this change in more detail. Table
\@ref(tab:dap-summary) presents an aggregate summary of individuals who kept and 
changed their plans. Individuals with no wheelchair users in their household are 
not affected, other than through some random simulation error. For wheelchair users
themselves however, the changes are relatively large. 
For example, `r change1*100` percent of the wheelchair users who were modeled as 
making mandatory tour patterns in the base scenario are now expected to make 
non-mandatory or stay-at-home patterns. Similarly, wheelchair users who previously
chose non-mandatory patttens are considerably more likely to stay at home.

The coordinated nature of the ActivitySim daily activity pattern model also allows
household members of wheelchair users to be affected. For these people, 
the changes are smaller in proportion but potentiall meaningful in aggregate: 
`r change2 * 100` percent of the household members who stay at home in the new scenario
incorporating wheelchair use previously made out-of-home tours in the base scenario.


```{r dap-map, fig.cap="Percent of households with changed daily activity pattern between two scenarios.", fig.height = 8, cache=FALSE}
tar_load(dap_map)
dm <- dap_map %>% filter(pct > 0)

if(!knitr::is_latex_output()){
  ncolor <- 5
  mfill <- colorNumeric(viridis(ncolor, direction = -1,  option = "inferno"), 
                        domain = dm$pct, n = ncolor)
  leaflet(dm) %>%
    addProviderTiles(providers$CartoDB) %>%
    addPolygons(fillColor = ~mfill(pct), stroke = 0, fillOpacity = 0.8) %>%
    addLegend(pal = mfill, values = ~pct, title = "P changed DAP") 

  
} else {
  ggplot(dm, aes(fill = pct)) +
    annotation_map_tile("cartolight", zoom = 10) +
    geom_sf(alpha = 0.5, color = NA) +
    scale_fill_viridis_c("Percent changed DAP", option = "inferno", direction = -1,
                         labels = scales::percent_format(accuracy = 5L, suffix = "" )) +
    theme(
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.text=element_text(size=8), #change font size of legend text
      legend.title=element_text(size=8)
    )
  
}

```


In addition to these aggregate numbers, it is also worth considering where the
affected households are concentrated. Wheelchair use and disability in general
is not distributed evenly through the region. Figure \@ref(fig:dap-map) shows the
concentration of households where at least one member changed its chosen 
daily activity pattern between the scenarios by traffic analysis zone. Though
most areas of the region experience little if any alteration, some zones see as 
many as 20 percent or 30 percent of their households change activity patterns. 












