# Applications

## Example one NHTS

We estimated the models using mlogit for R \cite{R, mlogit}. As described above, the
alternatives for daily activity pattern choice are a Mandatory pattern where the 
individual's day involves a work or school tour, a Non-Mandatory pattern where 
only discretionary trips are taken, and a Home pattern where the individual does 
not leave home. In the models estimated for this study, the Home pattern serves
as the reference alternative with a utility of zero. Retired and otherwise 
non-working individuals choose only between Non-Mandatory and Home daily activity
patterns.

The model estimates are presented in 
Table @ref(tab:model-coef). 
The estimated coefficients are of the expected sign, though not all are significant. 
Some predictors that proved to be insignificant, such as automobile availability
for full-time workers, were excluded from the estimated models. 
The overall model fit --- as indicated by the McFadden $\rho^2$ with respect to
a market shares (constants only) model --- is not strikingly high. Were the purpose
of this research to identify the best fit model of activity pattern choice for 
each person segment we would undertake an exercise to include, exclude, and identify
potential transformations for different sets of variables. In this case, however,
the goal of these models is simply to provide a plausible comparison point for
the behavior of individuals using wheelchairs against the behavior of individuals
in other person type segments. 

```{r model-coef}
tar_load(pt_modelsummary)
modelsummary(
  pt_modelsummary, group = variable + term ~ model, 
  estimate = "{estimate} ({statistic}){stars}",
  statistic = NULL,
  notes = c("t-statistics in parentheses, * p < 0.1, ** p < 0.05, *** p < 0.01",
            "Coefficients represent utility change relative to H: stay at home pattern."),
  title = "Daily Activity Pattern Model Estimates") %>%
  kable_styling(latex_options = c("scale_down"))
```


In this regard, the model results show strong divergence of the utility
preferences of individuals who use wheelchairs. For instance, full-time workers 
in the middle income groups are modestly less likely to choose non-mandatory 
patterns, and part-time workers of higher income are less like to choose
mandatory patterns. Income appears to have no discernible effect on  the choices
of non-working and retired individuals. We see a negative utility score for all 
person types with a wheelchair variable, and “mandatory” is even more negative. 
This is expected as individuals with wheelchairs are less likely to take a work 
or school trip compared to a shopping or a recreational trip. Non-workers and 
retired person types do not have a coefficient for “mandatory” DAP because those 
users by definition do not take “mandatory” DAP. Indeed, wheelchair use is among 
the strongest predictors of daily activity pattern choice across population 
segments. 

## Example two ActivitySim

As a secondary example, the research measures the impact of 
wheelchair status on ActivitySim’s selection of daily plans for our given 
synthetic population. Given a “Before” scenario in ActivitySim of the Salt Lake 
Area and ignoring the newly added wheelchair status in the synthetic population, 
ActivitySim predicted a DAP for each individual. In a second, “After” scenario, 
ActivitySim again predicted a DAP for each person, this time considering the 
wheelchair use status of each individual in the population. We hypothesized that 
those with wheelchairs and those in the same households as individuals with 
wheelchairs would change their DAP because of the negative utility scores 
applied to the “mandatory” and “nonmandatory” DAP alternatives, and the rest of 
the population would be unaffected. The DAP of those within the same household 
of a wheelchair user may change because of the coordinated nature of household 
DAP in ActivitySim. Table 4-6 shows the change in DAP among those with 
wheelchairs, in the same household as one with a wheelchair, and with neither a 
wheelchair nor in the same household. The table contains both total volumes and 
percentages; the value of percent is by total volume in the group, for example, 
16.4 percent of Wheelchair Users chose a “home” pattern in both the “Before” 
scenario and the “After” scenario. The latter group is rightly unaffected by 
the wheelchair implementation in the simulation (with the exception of a few
changes attributable to randomness) and does not include a percentage breakdown. 
Primarily, DAP remain the same for most individuals, as shown in the diagonal. 
However, there is a large volume of wheelchair users and their household members 
that stay home, particularly from “nonmandatory” DAP. This finding is consistent 
with our hypothesis. 

```{r dap-summary}
tar_load(dap_table)
kbl(dap_table, booktabs = TRUE, caption = "Daily Activity Pattern Change",
    col.names = c("Group", "DAP without Wheelchair Use", "H", "M", "N")) %>%
  add_header_above(c("", "", "DAP with Wheelchair Use" = 3)) %>%
  collapse_rows(columns = 1, latex_hline = "major") %>% 
  kable_styling()
```

















