---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# wheelchair_cdap2

<!-- badges: start -->
<!-- badges: end -->

This repository contains data analysis and manuscript files related to a study
on how overall daily activity patterns are affected by the use of a wheelchair.

## Analysis

The analysis can be executed with the `targets` package. To do this, clone the
repository, install `targets`, and then run the following commands:

```{r targets}
library(targets)
tar_make()
```

The analysis uses several R packages. Users can install the necessary 
packages by doing the following:

```{r packages, eval = FALSE}
# install.packages("devtools") # devtools: install non-CRAN pkgs on GitHub
devtools::install_github("byu-transpolab/apollo") # personal fork of apollo
devtools::install_github("byu-transpolab/nhts2017") # NHTS 2017 tables

install.packages("mlogit")
install.packages("tidyverse")
```


## Document

The document is written in `bookdown` using the targets created above.